---
title: "ESM 206 Lab 5"
author: "Audrey Nolan"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 objectives: 

- learn to parse dates with teh 'lubridate' package
- get counts of observations with counts
- uncount() rows 
- one and two-sample t-test
- create a heat map with geom_tile()

```{r, include = FALSE}
# Read in data and attach packages
library(tidyverse)
library(here)
library(janitor)

# read in data then wrangle 

lobster_abundance <- read_csv(here::here( "data", "lobster_abundance.csv"), 
                              na = "-99999") %>%
  janitor::clean_names()
  

```

Use tidyr:: uncount() function to convert our lobster data from frequency format to case format (give it the column name that contains the current counted info)

```{r}

lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)

```

#### Exploratory data visualization

- We're only going to consider 'site' as our variable of interest

```{r}

ggplot(lobster_tidy, aes( x = site, y = size_mm)) + 
  geom_jitter(aes(color = site), 
              alpha = 0.5, 
              width = 0.2)

# lets make a histogram with a group for each site ( facet wrap to sep based on site, other wise itll be an overwhelming stacked histogram ), scales = free tells it to make the scale of each graph dependent on that individual set, not useful if you are trying to compare the sets, can be misleading 


ggplot(lobster_tidy, aes(x = size_mm))+
  geom_histogram(aes(fill = site)) + 
  facet_wrap(~site, scales = "free")

# geomqq needs to know what the sample; looking for does there seem to be a normally distributed sample(linear)? ... does this make sense for this data to be linear within a means comparison? if no, then when means are norm, we can use a t-test (central limit theorem)

ggplot(lobster_tidy, aes(sample = size_mm))+
  geom_qq()+
  facet_wrap(~site)

```


#### Convert the 'date' columb to class 'Date'
 - using 'lubridate' so R knows that this is a Date format, and to help us easily parse month & year 
using mutate to the R what the old is MDDYY, so need to give it this existing order and which column name currently exists with that info 
 
```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date))

```
 
 
Now, lets parse year and month using lubridate::month( )(pulls just month number out of a number that is in date format, also automatically assumes it's an ordered factor and the order should be maintained) and lubridate::year()
 
```{r}

lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE), 
    obs_year = lubridate::year(date_new)
  )

```
 
Now, let's find counts of obeserved lobsters based on different heirarchical groupings: 

say we wanna know for each year, each month, what is total lobster count? 
```{r}

lobster_ym <- lobster_parse_date %>%
  dplyr::count(obs_year, obs_month)

lobster_y <- lobster_parse_date %>% 
  count(obs_year)

lobster_site <- lobster_parse_date %>% 
  count(site)

```

If we want to create a summary table that contains stats other than counts, it is easier to use group_by() + n()

```{r}

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    lobster_mean = mean(size_mm, na.rm = TRUE), 
    sd_size = sd(size_mm, na.rm = TRUE), 
    lobster_number = n())

```

#### Find Confidence intervals

Use t.test() function to find confidence intervals for one sample and perform t-test to compare means of two samples (covered in lectures from week 6)
use pull to pull a col from a df
```{r}

ivee_lobster <- lobster_tidy %>%
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobster)

```

#### Two-sample t-test to compare means
idk if correct-- > are my means diff enough, that i think that these samples were drawn from two diff populations that have different means?? 

We want to ask: Is there a significant difference in lobster lengths at Naples and Mojawk reefs?

Weve done our necessary exploratory analyses to determine that t 2-sample t-test for means comparison is appropriate. 

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample)

mn_ttest

```

There is a significant diff in lobster lengths b/t Naples and Mohawk Reef(t(df) = stat, p < 0.001, alpha = 0.05). 

does a t test as a func of the different levels within whatever data col you give it 
```{r}

lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(data = lobster_mn, size_mm ~ site)

mn_ttest2

mn_ttest2$p.value
mn_ttest2$statistic

```
use in-line variable referencing to call parts from specific tests !!!!!!! (mn_ttest2$statistic)

Here is the p-value: `r mn_ttest2$p.value`

Here is my t-stat : `r mn_ttest2$statistic`

#### Nowww a heat map:

```{r}

lobster_ys <- lobster_parse_date %>% 
  count(obs_year, site)


ggplot(data = lobster_ys, aes(x = obs_year, 
                              y = site)) +
  geom_tile(aes(fill = n))


```


